---
date: "2014-09-28"
tags:
- Hubway
- theme
- command-line
title: HUBWAY DATA EXERCISE
---

 
*** 
 
 
 
##1
### Network Visualisation : Node-link
#### Ride Network of Nodes 7, 8 and 9

The network has 3 sub networks of Nodes 7, 8 and 9 connected with their respective rides edges. A simple representation.


```{r, include=FALSE}
library(igraph)
library(tidyverse)
Hubway1 <- read.delim("Hubway Tej.txt", header = FALSE)
 


```



```{r}
smpl <- graph.data.frame(Hubway1)

igraph.options(layout=layout.kamada.kawai,
               vertex.label = NA,           
               vertex.size = 5,            
               vertex.frame.color = "blue",     
               vertex.color = "lightblue") 

plot(smpl, edge.arrow.size=0.1)

```



*** 

##2
### Network Visualisation : Node-link (Normalised)
#### Ride Network of Nodes 7, 8 and 9

Same as above, the network has 3 sub networks, with the normalised values. The size diffrence in Nodes 7, 8 and 9 is clearly visible, but it is little difficult to understand without the lables. Also the normalisation has transformed the edges into points. This can be helpful to understand the importance/ magnitude of the Nodes.


```{r}
V(smpl)$degree       <-  degree(smpl, normalized = TRUE)
V(smpl)$betweenness  <-  betweenness(smpl, normalized = TRUE)
V(smpl)$closeness    <-  closeness(smpl, normalized = TRUE)


igraph.options(layout=layout.kamada.kawai,
               vertex.label = NA,           
               vertex.size = 10,            
               vertex.frame.color = "darkblue",     
               vertex.color = "white")  

V(smpl)$size  <- V(smpl)$degree * 100   
V(smpl)$color <- V(smpl)$betweenness * ncol(smpl)  
V(smpl)$color <- V(smpl)$closeness   * ncol(smpl)    
plot(smpl, edge.arrow.size=0.3)


```
*** 
##3
### Network Visualisation : Node-link (Interactive)
#### Ride Network of Nodes 7 and another sample network

The interactive ride network of Node 7 - Fan Pier is connected with its rides edges. 
```{r}

library(networkD3)
Hubway2 <- read.delim("Hubway Tej Intractive.txt", header = FALSE)
n <- simpleNetwork(Hubway2)
n
```


#### Sample Network
To understand the network and the connections within, by interactive means.

```{r}

Hubway3 <- read.delim("Hubway Tej Intractive2.txt", header = FALSE)
simpleNetwork(Hubway3)


```

